---
title: "Lab1"
author: "Denys Kushnirenko"
date: "2023-09-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Calculate the measures of central tendency: mean, median.

```{r, echo=FALSE}
drug_deaths <- read.csv(".\\drug_deaths.csv")
```
```{r}
mean(drug_deaths$Age,na.rm=T)
median(drug_deaths$Age,na.rm=T)
```
## 2. Calculate the measures of variability: variance, standard deviation, coefficient of variation, range of variation and interquartile range.

```{r}
var(drug_deaths$Age,na.rm=T)
sd(drug_deaths$Age,na.rm=T)
sd(drug_deaths$Age,na.rm=T)/mean(drug_deaths$Age,na.rm=T)
max(drug_deaths$Age,na.rm=T)-min(drug_deaths$Age,na.rm=T)
IQR(drug_deaths$Age,na.rm=T)
```

## 3. Create Box and Whisker plot

```{r}
boxplot(drug_deaths$Age, ylab = "Age", main = "Box and Whisker for Age variable")
```

## 4. Output the five-point summary of a sample (extreme points and quartiles).

```{r}
summary(drug_deaths$Age,na.rm=T)
```

## 5. Find the 1st and 9th deciles (without intermediate).

```{r}
decile <- quantile(drug_deaths$Age, seq(0,1,.1),na.rm=T)
cat("first decile =",decile[2],"\nnineth decile =",decile[10])
```

## 6. Calculate the skewness coefficient and the excess coefficient.

```{r}
library(EnvStats)
skewness(drug_deaths$Age,na.rm=T) 
kurtosis(drug_deaths$Age,na.rm=T)
```

## 7. Construct a histogram using different binning methods (basic rule, Scott's rule, Friedman-Diaconis rule). On the same graph, plot the density estimate and the density plot of the hypothesized distribution. Formulate assumptions about the type of distribution observed.

```{r,echo=FALSE}
library(gamlss)
```
```{r}
hist(drug_deaths$Age, freq=F, xlab = "Age", ylab='Relative frequency',
     main="Sturges rule", col = "blue")
hist(drug_deaths$Age, breaks = "Scott", freq=F, xlab = "Age",
     ylab='Relative frequency', main="Scott's rule", col = "red")
hist(drug_deaths$Age, breaks = "freedman-diaconis", freq=F, xlab = "Age",
     ylab='Relative frequency', main="Friedman-Diaconis rule", col = "darkgreen")
histDist(na.omit(drug_deaths$Age), family = "NO", density = TRUE)
lines(density(na.omit(drug_deaths$Age), na.rm = T), col='blue', lwd=1)
```

## 8. Plot a Q-Q plot to check for agreement with the hypothesized distribution.

```{r}
qqnorm(drug_deaths$Age, col="darkblue")
qqline(drug_deaths$Age, lwd = 2)
```

## 9. Plot a P-P plot to check for agreement with the hypothesized distribution.

```{r,echo=FALSE}
library(ggplot2)
library(qqplotr)
library(nortest)
```
```{r}
dp <- list(mean = 41.96, sd = 12.34)
ggplot(data = drug_deaths, mapping = aes(sample = Age)) +
  stat_pp_band(dparams = dp) +
  stat_pp_line() +
  stat_pp_point(dparams = dp) +
  labs(x = "Probability Points", y = "Cumulative Probability")
```

## 10. Use one of the statistical tests to check for agreement with the hypothesized distribution.
```{r}
lillie.test(drug_deaths$Age)
```